---
title: "Spécific K-mer selection for RNA virus VS mosquito host"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include = FALSE}
library(Biostrings)
library(stringr)
library(dplyr)
library(ggplot2)
library(KmerEnrich)
library(data.table)
library(ShortRead)
```

# Parameters for the Pipeline

```{r define pipeline variables}
virus_genome_path <- "../data_processed/virus/dengue/dengue_type_1/"
vector_genome_path <- c("../data_processed/vector/aedes/aegypty/SRR29420355",
                        "../data_processed/vector/aedes/aegypty/SRR23079315")

k <- 6
x <- 100
```

# Pipeline

Generation of the list of K-mers of interest based on the kmers present in all sequences of "dengue_virus_type_1" that are only appearing in the first x = 100 nulceotides of size k = 6

```{r Generate k-mers list}
best_kmers_list <- common_kmers(virus_genome_path,k,x)
```

Generation of the information reguarding the position of every kmer in all sequences, the distances between the kmers and between the start position and the end position.

Then using the get_stat_pos_df function and implementing the cleated dataframe inside of the v_kmer_pos_df function this will generate the statistical information about all kmers, their frequence, their repartition on the genomes the amount of times they appear on average.

```{r Get info reguarding specific k-mers reguarding the virus}
v_kmer_pos_df <- kmers_pos_df(virus_genome_path,best_kmers_list)

v_kmers_stats <- get_virus_stats_kmers(v_kmer_pos_df)
```

Using the previoulsy genereted dataframe containing all information about the viruses we use the function get_kmer_freq to get the frequency of apparition of the kmers in the RNA sequences of the vectors(mosquitos) completing the dataframe with the background information.

```{r Get statistical information of k-mer reguarding the mosquito}
complete_kmer_stats <- get_vector_stats_kmers(v_kmers_stats, vector_genome_path)
```

Using all the statisctical values use them to generate a plot alowing us to represent the best Kmers and select the one that is desiered reguardin the scenario you need. The X axis will show us how well distributed the kmer is in the virus genome The Y axis will represents the ratio of how much times the kmer appears more on the virus genome then on average in the RNAseq values of the mosquito.

```{r Plot the data}
complete_kmer_stats <- complete_kmer_stats %>%
  filter(mean_count > 5)

ggplot(complete_kmer_stats, aes(x = Distribution_score, y = Enrichment_score, color = accession_id)) +
  geom_line(aes(group = Kmer), color = "grey", linetype = "dashed") +
  geom_point(shape = 16, size = 2) +
  geom_text(aes(label = paste0(Kmer, "\n", mean_count)), 
            vjust = -0.5, size = 2, show.legend = FALSE) +
  labs(title = "frequency_virus / frequency_mos",
       x = "mean count per virus / max log10 segment in virus genome",
       y = " frequency virus / frequency mos") +
  theme_minimal() +
  theme(axis.text.x = element_text(hjust = 1)) +
  expand_limits(x = max(complete_kmer_stats$Distribution_score) * 1.05, 
                y = max(complete_kmer_stats$Enrichment_score) * 1.1)

```
